---
title: "flat_preprocess_2018.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

# preprocess_2018

In this notebook, we're exploring and preprocessing the 2018 data from OpenSRP.
The function will ultimately preprocess the data so that it can be harmonized
effectively with the other datasets.

What is OpenSRP?

What is OpenSRP? It appears OpenSRP should cover year 2018. What is it and what does it survey?
> Open SRP is a software to collect data. We collaborated with an Indonesian-based researchin 2018 that managed and stored our data. Their website does not exist anymore. That is the first platform we used for the ongoing health surveillance surveys. The data is stored in a flat file format. We are working on a new platform to collect data, but we are not there yet. The data is stored in a flat file format. We are working on a new platform to collect data, but we are not there yet.

## EDA

Let's start with some simple EDA.

```{r development-preprocess_2018}
# Prepare the code of your function here
library(here)
library(dplyr)
library(stringr)
library(janitor)
library(targets)
library(skimr)
library(googledrive)

if(!authenticate_google_drive()) {
  stop("Google drive not authenticated!")
}
```

```{r development-fetch_data}
tar_load(opensrp)
tar_load(opensrp_dict)
```

The first thing to do is to run the column names through `clean_names`:

```{r}
opensrp_ <- clean_names(opensrp)
```

Now, we go through each column and check the types and summary of the data:

```{r}
summary(opensrp_)
```

```{r}
skim(opensrp_)
```

What is user ID?

```{r}
opensrp_ %>%
  pull(user_id) %>%
  unique() %>% length()
```

We can leave this as a factor since it defines

```{r function-preprocess_2018}
#' preprocess_2018 Title
#'
#' @return 1
#' @export
#'
#' @examples
preprocess_2018 <- function() {
  1
}
```

```{r examples-preprocess_2018}
preprocess_2018()
```

```{r tests-preprocess_2018}
test_that("preprocess_2018 works", {
  expect_true(inherits(preprocess_2018, "function"))
})
```

### 2018

To start, we read in the 2018 data:

```{r}
library(dplyr)
```

```{r}
drive_find(pattern = "openSRP_data_2018*") %>%
  select(-id)
```

Since it's an xlsx, we'll need to use the appropriate library:

```{r}
library(readxl)
library(here)
```

```{r}
drive_find(pattern = "openSRP_data_2018*") %>%
  drive_download(path = here("data", "raw", "openSRP_data_2018.xlsx"), overwrite=TRUE) -> dat

openSRP <- read_excel(dat$local_path)

#skimr is a great way to look at the data without
#having to print the whole thing and reveal PHI
skimr::skim(openSRP)
```

There are loads of columns:

```{r}
names(openSRP)
```

I think using `janitor` across this project will help with aligning column
types and names.

```{r}
library(janitor)
clean_names(openSRP) %>%
  skimr::skim()
```

We also have a data dictionary available:

```{r}
drive_find(pattern = "openSRP_dictionary*") %>%
  drive_download(path = here("data", "raw", "openSRP_dictionary.xlsx"), overwrite=TRUE) -> dat

openSRP_dict <- read_excel(dat$local_path)

head(openSRP_dict)
```

#### Column by Column

Let's investigate all of the columns in the dataset.



Let's start with addressing column types:

```{r}
openSRP_ <- clean_names(openSRP)
```

```{r}
numerics <- c("", "")
factors <- c("")
strings <- c("")
```

```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_preprocess_2018.Rmd", vignette_name = "Go further")
```

