---
title: "flat_preprocess_2018_health.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r setup}  
library(here)
library(targets)
tar_load(opensrp, store = here::here("_targets"))
tar_load(opensrp_dict, store = here::here("_targets"))
tar_load(cohort_2018_deid, store = here::here("_targets"))

# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)

library(janitor)
library(dplyr)
library(tidyr)
library(lubridate)
library(skimr)
library(forcats)
library(stringr)
```

In this notebook we're tackling the health data from the cohort 2018 opensrp file. We decided to
leave this to a separate processing step because the health data is a bit more complex than the household registration and demographic data.

Let's begin:

## Reading in the data

The data is read in from the targets pipeline, and is contained in the `opensrp` object.

```{r}
open_census <- opensrp$`Open census`

head(open_census)
```

In this _specific case_, the raw data in the open census file is the same number of rows as the
processed targets output, so I'm going to join the raw data to the health columns.

```{r}
open_census %>%
  clean_names() %>%
  # remove this individual as they are specifically excluded from the cohort as a duplicate
  filter(!str_detect(member, "36841201")) %>%
  select(prior_health_care:deathyear) %>%
  bind_cols(select(cohort_2018_deid, uuid_clean = uuid, sex_clean = sex, age_group_clean = age_group), .) -> health_data
```

Now let's take a look at the data:
```{r}
skimr::skim(health_data)
```

## Health Care Access & Diagnoses

The first thing to determine is whether `prior_health_care` refers to,
"does this person have a history of a health care condition," or something else.

```{r}
health_data %>%
  select(prior_health_care, prior_diagnosis) %>%
  table()
```

From this it would appear that prior healthcare is a yes/no question
that tells us if they have any history of health care conditions.

```{r}
health_data %>%
  mutate(prior_health_condition_clean = case_when(
    prior_health_care == "Yes" ~ "Yes",
    !is.na(prior_diagnosis) ~ "Yes",
    TRUE ~ "No"
  )) -> health_data_
```

Now let's look at the `prior_diagnosis` column:

```{r}
health_data_ %>%
  select(prior_diagnosis) %>%
  table()
```

```{r}
health_data_ %>%
  select(other_prior_diagnosis) %>%
  table()
```

In the secondary column, we see many values that are translated. We'll translate
these manually and then cast them width wise to create a wide format table.

First, we mutate the `prior_diagnosis` column to make sure a diagnosis is one word, or clearly separated by a comma:

```{r}
health_data_ %>%
  mutate(
    prior_diagnosis_ =
      prior_diagnosis %>%
        str_to_lower() %>%
        str_replace("infected wound prior diagnosis other", "infected_wound") %>%
        str_replace("malaria cancer", "malaria,cancer") %>%
        str_replace("malaria prior diagnosis other", "malaria") %>%
        str_replace("no pregnancy prior diagnosis other", "no_pregnancy") %>%
        str_replace("no pregnancy", "no_pregnancy") %>%
        str_replace("prior diagnosis other", "other") %>%
        str_replace("hypertension idr", "hypertension") %>%
        str_replace("no diagnosis", NA_character_) %>%
        str_replace("idr", NA_character_)
  ) %>%
  select(prior_diagnosis_) %>%
  table()
```

Then, we concatenate the `prior_diagnosis` and `other_prior_diagnosis` columns to create a single diagnosis column separated by commas:

```{r}
health_data_ %>%
  mutate(
    prior_diagnosis_ =
      prior_diagnosis %>%
        str_to_lower() %>%
        str_replace("infected wound prior diagnosis other", "infected_wound") %>%
        str_replace("malaria cancer", "malaria,cancer") %>%
        str_replace("malaria prior diagnosis other", "malaria") %>%
        str_replace("no pregnancy prior diagnosis other", "no_pregnancy") %>%
        str_replace("no pregnancy", "no_pregnancy") %>%
        str_replace("prior diagnosis other", "other") %>%
        str_replace("hypertension idr", "hypertension") %>%
        str_replace("no diagnosis", NA_character_) %>%
        str_replace("idr", NA_character_)
  ) %>%
  mutate(
    other_prior_diagnosis =
      other_prior_diagnosis %>%
        str_to_lower() %>%
        str_replace("infected wound prior diagnosis other", "infected_wound") %>%
        str_replace("malaria cancer", "malaria,cancer") %>%
        str_replace("malaria prior diagnosis other", "malaria") %>%
        str_replace("no pregnancy prior diagnosis other", "no_pregnancy") %>%
        str_replace("no pregnancy", "no_pregnancy") %>%
        str_replace("prior diagnosis other", "other") %>%
        str_replace("hypertension idr", "hypertension") %>%
        str_replace("no diagnosis", NA_character_) %>%
        str_replace("idr", NA_character_)
  ) %>%
  unite(diagnoses_all, prior_diagnosis_, other_prior_diagnosis, sep = ",", remove = FALSE, na.rm = TRUE) %>%
  mutate(
    diagnoses_all =
      str_squish(diagnoses_all) %>%
        str_replace(", ", ",") %>%
        str_replace(" ", "_") %>%
        # if it just starts with "other," it means it has only "other diagnosis"
        # and can be safely removed
        str_replace("^other,", "") %>%
        # which means the ones that are 999 are true "other" diagnoses
        # with no specification in english
        str_replace("999", "other_unspecified")
  ) %>%
  mutate(diagnoses_all = na_if(diagnoses_all, "")) %>%
  select(diagnoses_all) %>%
  table()
```

Now we have a single column with all diagnoses separated by a comma. We can
use this to create a wide format table with each diagnosis as a column.

These values can then be translated and categorised into ICD10 categories, which might be useful
for cross-cultural comparisons and publications.
    
```{r malagasy_diagnosis_lookup_table}
malagasy_diagnosis_lookup_table <- tibble::tribble(
  ~raw, ~diagnosis_english, ~icd10_code, ~icd10_description,
  "aretim_bavony", "stomach_ache", "R10–R19", "Symptoms and signs involving the digestive system and abdomen",
  "adetin_kibo", "diarrhea", "R10–R19", "Symptoms and signs involving the digestive system and abdomen",
  "farasisa", "heel_fissures", "L80–L99", "Other disorders of the skin and subcutaneous tissue",
  "hernie", "hernia_prolapse", "K40–K46", "Hernia",
  "voatombo_komby", "injury_by_animal_zebu", "W50–W64", "Exposure to animate mechanical forces",
  "zonisy", "jaundice", "R10–R19", "Symptoms and signs involving the digestive system and abdomen",
  "aretintsofina", "ear_nose_throat", "H60–H95", "Diseases of the ear and mastoid process",
  "hoditra_fotsy", "genetic_albinism", "E70–E90", "Metabolic disorders",
  "solopiso", "syphilis_sti", "A50–A64", "Infections with a predominantly sexual mode of transmission",
  "mentally_handicapped", "developmental_cognitive", "F70–F79", "Intellectual disabilities",
  "bongabe", "big_testicles", "N40–N51", "Diseases of male genital organs",
  "mangoraka", "shivering_trembling", "R20–R29", "Symptoms and signs involving the nervous and musculoskeletal systems",
  "tadigny", "ear_ache", "H60–H95", "Diseases of the ear and mastoid process",
  "matetika_tontona", "epistaxis_nosebleed", "R00–R09", "Symptoms and signs involving the circulatory and respiratory systems",
  "tontona", "epistaxis_nosebleed", "R00–R09", "Symptoms and signs involving the circulatory and respiratory systems",
  "sohiky", "asthma", "J40–J47", "Chronic lower respiratory diseases",
  "soiky", "asthma", "J40–J47", "Chronic lower respiratory diseases",
  "torantorana", "dizziness_syncope", "R40–R46", "Symptoms and signs involving cognition, perception, emotional state and behavior",
  "hypertension", "hypertension", "I10–I15", "Hypertensive diseases",
  "malaria", "malaria", "B50–B64", "Protozoal diseases",
  "tuberculosis", "tuberculosis", "A15–A19", "Tuberculosis",
  "infected_wound", "skin_soft_tissue_infection", "L00–L08", "Infections of the skin and subcutaneous tissue",
  "no_pregnancy", "pregnancy_related", "O00–O99", "Pregnancy, childbirth and the puerperium",
  "cancer", "neoplasm_unspecified", "C00–D48", "Neoplasms",
  "other", "other_unknown", "R00–R99", "Symptoms, signs and abnormal clinical and laboratory findings, not elsewhere classified"
)
# Save to package
usethis::use_data(malagasy_diagnosis_lookup_table, overwrite = TRUE)
```

You can examine this lookup table in the vignette [`Pipeline Outputs`](OutputTables.html).

```{r function-malagasy_diagnosis_lookup_table}
#' Malagasy Diagnosis Lookup Table
#'
#' A lookup table mapping raw diagnosis entries (from Malagasy or English inputs)
#' to standardized English diagnosis labels and ICD-10 chapter-level codes.
#'
#' Used in conjunction with `create_health_diagnoses()` to clean and harmonize
#' multi-source survey data (e.g., `prior_diagnosis`, `other_prior_diagnosis`, or `diagnoses_all`).
#'
#' @format A tibble with 3 columns:
#' \describe{
#'   \item{raw}{Raw input string from survey data}
#'   \item{diagnosis_english}{Cleaned harmonized English label}
#'   \item{icd10_code}{ICD-10 chapter or block code (e.g. "K00–K95")}
#' }
#' @source Manually curated from OpenSRP 2018 health survey entries in the \link[=preprocessing-2018-health-status]{Health Preprocessing} vignette.
"malagasy_diagnosis_lookup_table"
```
  
With that, we can implement it in the function:

```{r development-create_health_diagnoses}
# You can prepare the code of the create_health_diagnoses() function here
```
  
```{r function-create_health_diagnoses}
#' Create Harmonized Health Diagnosis Variables
#'
#' This function takes a dataframe with a column of raw health diagnoses (`prior_diagnosis`)
#' in Malagasy or local terms, and returns a modified dataframe with three new variables:
#' - `diagnosis_english`: a harmonized, readable label in English
#' - `diagnosis_icd10_category`: a corresponding broad ICD-10 category name
#' - `diagnosis_icd10_code`: the ICD-10 code range (e.g., "K00–K95")
#'
#' The mappings are based on field-specific translations and align with global health reporting
#' frameworks such as DHIS2 and ICD-10 Chapter-level groupings.
#'
#' @param df A dataframe containing a column `prior_diagnosis` with free-text or categorical health terms.
#'
#' @return A dataframe with additional columns: `diagnosis_english`, `diagnosis_icd10_category`, and `diagnosis_icd10_code`.
#'
#'
#' @export
create_health_diagnoses <- function(df, diagnosis_lookup = malagasy_diagnosis_lookup_table) {
  df %>%
    mutate(
      prior_diagnosis_ =
        prior_diagnosis %>%
          str_to_lower() %>%
          str_replace("infected wound prior diagnosis other", "infected_wound") %>%
          str_replace("malaria cancer", "malaria,cancer") %>%
          str_replace("malaria prior diagnosis other", "malaria") %>%
          str_replace("no pregnancy prior diagnosis other", "no_pregnancy") %>%
          str_replace("no pregnancy", "no_pregnancy") %>%
          str_replace("prior diagnosis other", "other") %>%
          str_replace("hypertension idr", "hypertension") %>%
          str_replace("no diagnosis", NA_character_) %>%
          str_replace("idr", NA_character_)
    ) %>%
    mutate(
      other_prior_diagnosis =
        other_prior_diagnosis %>%
          str_to_lower() %>%
          str_replace("infected wound prior diagnosis other", "infected_wound") %>%
          str_replace("malaria cancer", "malaria,cancer") %>%
          str_replace("malaria prior diagnosis other", "malaria") %>%
          str_replace("no pregnancy prior diagnosis other", "no_pregnancy") %>%
          str_replace("no pregnancy", "no_pregnancy") %>%
          str_replace("prior diagnosis other", "other") %>%
          str_replace("hypertension idr", "hypertension") %>%
          str_replace("no diagnosis", NA_character_) %>%
          str_replace("idr", NA_character_)
    ) %>%
    unite(diagnoses_all, prior_diagnosis_, other_prior_diagnosis, sep = ",", remove = FALSE, na.rm = TRUE) %>%
    mutate(
      diagnoses_all =
        str_squish(diagnoses_all) %>%
          str_replace(", ", ",") %>%
          str_replace(" ", "_") %>%
          # if it just starts with "other," it means it has only "other diagnosis"
          # and can be safely removed
          str_replace("^other,", "") %>%
          # which means the ones that are 999 are true "other" diagnoses
          # with no specification in english
          str_replace("999", "other_unspecified")
    ) %>%
    mutate(diagnoses_all = na_if(diagnoses_all, "")) %>%
    # step 1: Unnest and match diagnoses
    mutate(id = row_number()) %>%
    separate_rows(diagnoses_all, sep = ",") %>%
    mutate(diagnoses_all = str_trim(diagnoses_all)) %>%
    left_join(diagnosis_lookup, by = c("diagnoses_all" = "raw")) %>%
    filter(!is.na(diagnosis_english)) -> long_df

  # Step 2: Create wide one-hot matrix
  wide_matrix <- long_df %>%
    mutate(present = TRUE) %>%
    distinct(id, diagnosis_english, present) %>%
    mutate(diagnosis_english = str_c("diag_", diagnosis_english, "_clean")) %>%
    pivot_wider(names_from = diagnosis_english, values_from = present, values_fill = FALSE)

  # Step 3: Join back to original data
  output <- df %>%
    mutate(id = row_number()) %>%
    left_join(
      long_df %>%
        group_by(id) %>%
        summarise(
          diagnosis_english_clean = paste(unique(diagnosis_english), collapse = ", "),
          diagnosis_icd10_code_clean = paste(unique(icd10_code), collapse = ", ")
        ),
      by = "id"
    ) %>%
    left_join(wide_matrix, by = "id") %>%
    select(-id)

  return(output)
}
```
  
```{r example-create_health_diagnoses}
health_data_ %>%
  create_health_diagnoses() -> health_data_

health_data_ %>%
  summary()
```
  
```{r tests-create_health_diagnoses}
test_that("create_health_diagnoses works", {
  expect_true(inherits(create_health_diagnoses, "function"))
})
```

## Vaccinations

We're going to categorize the vaccinations next:

```{r}
health_data_ %>% select(vaccine_card, vaccination_history) %>% table()
```

It looks like for the most part, when people have a vaccination card they have a history and can remember
which vaccinations they have had. I'm assuming Idk refers to folks who do not know _where_ their vaccination card
is, because these people can still confidently answer that they have had a specific vaccine.

For the folks who say No, it must mean that they do not have a vaccination card. However, there are some cases 
where people have a vaccination card but can say what vaccinations they have. Further, some can claim to have other
vaccinations besides those listed in the survey. 

```{r}
health_data_ %>% select(vaccine_card, vaccine_other) %>% table()
```

We can ignore this column as it is effectively the same as the vaccine history column. 
    
```{r development-clean_2018_vaccinations}
# You can prepare the code of the clean_2018_vaccinations() function here
```
  
```{r function-clean_2018_vaccinations}
#' Title
#' 
#' Description
#' 
#' @return
#' 
#' @export
clean_2018_vaccinations <- function(df){

  df %>%
    mutate(
      vaccine_card_clean = case_when(
        vaccine_card == "Yes" ~ "Yes",
        vaccine_card == "No" ~ "No",
        vaccine_card == "Idk" ~ "Unknown_or_misplaced",
        TRUE ~ NA_character_
      ),
      vaccination_history = 
        str_to_lower(vaccination_history) %>%
        str_replace("vacc idr", "other") %>%
        str_replace("vacc other", "other"),
      id = row_number()
    ) %>%
    separate_longer_delim(
      vaccination_history,
      delim = " "
    ) %>%
    mutate(vaccination_history = str_c("vacc_", vaccination_history, "_clean")) %>%
    mutate(flag = TRUE) %>%
    distinct(id, vaccination_history, .keep_all = TRUE) %>%
    filter(!is.na(vaccination_history)) %>%
    pivot_wider(
      names_from = vaccination_history,
      values_from = flag,
      values_fill = FALSE
    ) %>%
    select(-id)

}
```
  
```{r example-clean_2018_vaccinations}
health_data_ %>%
  clean_2018_vaccinations() -> health_data_
  
health_data_ %>%
  select(contains("vacc_")) %>%
  skimr::skim()
```
  
```{r tests-clean_2018_vaccinations}
test_that("clean_2018_vaccinations works", {
  expect_true(inherits(clean_2018_vaccinations, "function")) 
})
```
  
## Pregnancy

Next, pregnancy:

```{r}
health_data_ %>%
  select(contains("preg")) %>%
  table()
```

This will be fairly simple and we will just create a column that indicates whether the person
is currently pregnant or not, and how many pregnancies they've had in the past.

```{r}
health_data_ %>%
  select(number_deliveries, number_live_births) %>%
  table()
```

```{r}
health_data_ %>%
  select(birthweight, birthweight_gram) %>%
  mutate(across(everything(), as.numeric)) %>%
  summary()
```

There's not much to go off of here, but I'm assuming `birthweight_gram` is the weight of the most recent child.
    
```{r development-clean_2018_pregnancy}
# You can prepare the code of the clean_2018_pregnancy() function here
```
  
```{r function-clean_2018_pregnancy}
#' Title
#' 
#' Description
#' 
#' @return
#' 
#' @export
clean_2018_pregnancy <- function(df){

  df %>%
    mutate(
      pregnancy_status_clean = case_when(
        pregnant == "Yes" ~ "Currently_pregnant",
        pregnant == "No" ~ "Not_pregnant",
        TRUE ~ NA_character_
      ) %>% as.factor(),
      number_deliveries_clean = as.numeric(number_deliveries),
      number_live_births_clean = as.numeric(number_live_births),
      most_recent_birthweight_gram_clean = as.numeric(birthweight_gram)
    )
}
```
  
```{r example-clean_2018_pregnancy}
health_data_ %>% 
  clean_2018_pregnancy() -> health_data_

health_data_ %>%
  select(matches("pregnancy|birth|deliveries")) %>%
  select(contains("clean")) %>%
  skimr::skim()
```
  
```{r tests-clean_2018_pregnancy}
test_that("clean_2018_pregnancy works", {
  expect_true(inherits(clean_2018_pregnancy, "function")) 
})
```

## Other

There is no data in the columns dealing with deceased survey participants but we'll keep them
in for future comparison.

Looking at the surgical history:

```{r}
health_data %>%
  select(contains("surgeries")) %>%
  table()
```

These will likely need a deliberate translation.

Lastly, deformities will also get a translation:

```{r}
health_data %>%
  select(contains("deformit")) %>%
  table()
```

# preprocess_2018_health

The final function will combine all of these steps into a single function that can be used to preprocess the 2018 health data in the pipeline.

```{r development-preprocess_2018_health}
# Prepare the code of your function here
```

```{r function-preprocess_2018_health}
#' preprocess_2018_health Title
#'
#' @return 1
#' @export
#'
#' @examples
preprocess_2018_health <- function(open_census, cohort_2018_deid, diagnosis_lookup = malagasy_diagnosis_lookup_table) {
  
  open_census %>%
    clean_names() %>%
    filter(!str_detect(member, "36841201")) %>%
    select(prior_health_care:deathyear) %>%
    bind_cols(select(cohort_2018_deid, uuid_clean = uuid, sex_clean = sex, age_group_clean = age_group), .) -> health_data

  health_data %>%
    create_health_diagnoses(., diagnosis_lookup = diagnosis_lookup) %>%
    clean_2018_vaccinations() %>%
    clean_2018_pregnancy() %>%
    mutate(
      surgical_history_clean = case_when(
        surgeries == "Yes" ~ TRUE,
        surgeries == "No" ~ FALSE,
        TRUE ~ NA
      ),
      surgery_description_clean = case_when(
        # values taken from the table() of the prior_surgeries column
        is.na(prior_surgeries)                                              ~ NA_character_,
        prior_surgeries == "&"                                              ~ NA_character_,
        str_detect(prior_surgeries, "Niterak.*nididiana|Niterak nodidiana") ~ "Cesarean section",
        str_detect(prior_surgeries, "apandisite")                           ~ "Appendectomy",
        str_detect(prior_surgeries, "Voakiso.*tanan")                       ~ "Injury - hand trauma",
        str_detect(prior_surgeries, "famaky.*tongotra|lasety.*tongotra")    ~ "Injury - foot laceration",
        str_detect(prior_surgeries, "kakazo.*kirandra")                     ~ "Injury - ankle trauma",
        str_detect(prior_surgeries, "omb.*kibo")                            ~ "Injury - animal trauma (zebu)",
        TRUE ~ "Other / Unknown"
      ),
      deformities_clean = case_when(
        is.na(visual_health_deformities)              ~ NA_character_,
        visual_health_deformities == "Deformities na" ~ NA_character_,
        !is.na(visual_health_deformities) & !is.na(other_visual_health_deformities) ~ str_c(visual_health_deformities, ", ", other_visual_health_deformities),
        TRUE ~ visual_health_deformities
      )
    ) %>%
    select(contains("clean")) %>%
    rename_with(
      ~ str_replace(., "_clean$", ""),
      everything()
    )
  
}
```

```{r examples-preprocess_2018_health}
preprocess_2018_health(open_census, cohort_2018_deid) -> health_data_

skim(health_data_)
```

```{r tests-preprocess_2018_health}
test_that("preprocess_2018_health works", {
  expect_true(inherits(preprocess_2018_health, "function"))
})
```


```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_preprocess_2018_health.Rmd", vignette_name = "Preprocessing 2018 Health Status", check = FALSE, open_vignette = FALSE)
```

