---
title: "flat_preprocess_2019.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

# preprocess_srp2019

Similar to the previous notebook, we are goign to preprocess the 2019 dharma data in this notebook.
The goal is to create a function that will take the raw data and return a preprocessed version of the cohort
to merge with the 2018 cohort. As before, we use the column suffix `_clean`
to denote that the column has been cleaned and is being kept for analysis.

```{r}
library(here)
library(dplyr)
library(stringr)
library(janitor)
library(targets)
library(skimr)
library(forcats)
library(lubridate)
library(googledrive)
library(ggplot2)
library(ggalluvial)
```

## Data Structures

The data for the dharma 2019 collection is stored in an excel file with multiple sheets. As before,
we use the googledriver package to monitor and download it for access. In our targets pipeline,
this file was called `Dharma_followup_2020plus.xlsx` and was tracked 

First, we are going to load the data and do some basic EDA to see how the dharma data collection differs from
the opensrp data collection.

```{r}
tar_load(dharma2019)
```

Like the 2018 data, the dharma 2019 data is stored in a list of data frames corresponding to sheets in the excel.

> Level 0 Name is the household information such as the head of household and the GPS associated.
> Level 0 Coded is the same as the household information but the values of the variables are only coded.
> Level 1 Name is the individual member information. 
> Level 1 Coded (same comment as above for Level 0 Coded)

To find the cohort, we have to investigate the Level 1 Name sheet.

```{r}
dharma2019_dict <- dharma2019$`Data Dictionary`
names(dharma2019)
```

We know that our data dictionary is in the last sheet, so we can keep that in mind for when
we get to the involved questions. For now, let's try to recreate our cohort and see if they reproduce.

```{r}
level1_df <- dharma2019$`Level 1 Name`
```

```{r function-preprocess_2019}
#' preprocess_2019 Title
#'
#' @return 1
#' @export
#'
#' @examples
preprocess_2019 <- function() {
  1
}
```

```{r examples-preprocess_2019}
preprocess_2019()
```

```{r tests-preprocess_2019}
test_that("preprocess_2019 works", {
  expect_true(inherits(preprocess_2019, "function"))
})
```


```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_preprocess_2019.Rmd", vignette_name = "Go further")
```

