---
title: "flat_cohort_household_heads.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r setup}
library(here)
library(targets)
tar_load(opensrp, store = here::here("_targets"))
tar_load(opensrp_dict, store = here::here("_targets"))
tar_load(cohort_2018, store = here::here("_targets"))

# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)

library(janitor)
library(dplyr)
library(tidyr)
library(lubridate)
library(skimr)
library(forcats)
library(stringr)
```

# cohort_household_heads

In this stage of the pipeline, we're preprocessing and cleaning the dataset
to generate a record of the heads of households and their geographic locations.
This is important because it allows us to analyze the data at the household level,
which is often more relevant for understanding social and economic dynamics.

> In Madagascar, as in many countries, the household head (or chef de ménage) is a key unit of reference in population-level analyses, especially in surveys, censuses, and administrative datasets. In Madagascar, household headship can also reflect traditional norms and ethnic practices, which might influence the structure and function of households. Extended family households with multiple generations under one roof are common in some regions, and the definition of “head” may differ from a purely economic or legal interpretation.

This is therefore an important component of the data preparation project.

## Reading in the data

The data is read in from the targets pipeline, and is contained in the `opensrp` object.

```{r}
names(opensrp)
```

The specific sheet is called `Hhregistration`:

```{r}
hhreg <- opensrp$Hhregistration
head(hhreg)
```

## Cleaning the data

As in the 2018 cohort prep, we're going to start by cleaning the column names:

```{r}
hhreg %>%
  janitor::clean_names() -> hhreg_
```

```{r}
names(hhreg_)
```

```{r}
skim(hhreg_)
```

It looks like we will eventually drop the `sub_village` and hh_gps_point columns, as they contain only NA values.
In any case, GPS will be handled in a different file.

Is `user_id` anything useful?

```{r}
hhreg_ %>%
  select(user_id) %>%
  unique()
```

```{r}
hhreg_ %>%
  select(village) %>%
  unique()
```

How does that differ from the province?:

```{r}
hhreg_ %>%
  select(province) %>%
  unique()
```

Let's just look at these columns together:

```{r}
hhreg_ %>%
  select(user_id, village, sub_village, district, sub_district) %>%
  mutate(across(everything(), as_factor)) %>%
  summary()
```

So the only thing that changes is the use of Antaravato 1 and 2.

Are all of the heads in the cohort file represented in this file?

```{r}
cohort_2018 %>% 
  select(hh_head_name) %>% 
  unique() %>%
  pull() -> cohort_2018_hh_head_names

hhreg_ %>%
  mutate(hh_head_name = str_to_lower(name_household_head) %>% 
    str_to_lower() %>%
    str_squish() %>%
    str_remove_all(",") ) %>%
  select(hh_head_name) %>%
  unique() %>%
  pull() -> hhreg_hh_head_names

all(cohort_2018_hh_head_names %in% hhreg_hh_head_names)
```

As long as we follow the same preprocessing of names in the 2018 approach, the names all match.

## Adding Geography

We're also going to take care of cleaning the geography in this step.

```{r}
hh_geography <- opensrp$GPS
head(hh_geography)
```

There are two notes we'll have to resolve:

```{r}
hh_geography$Note %>% unique()
```

We do notice that there is an Eddy (I) and an Eddy (II) in the data.

We'll quickly clean this and join it into the household head dataframe:

```{r}
hh_geography %>%
  janitor::clean_names() %>%
  mutate(hh_head_name = str_to_lower(head_of_household) %>% 
    str_to_lower() %>%
    str_squish() %>%
    str_remove_all("\\(") %>%
    str_remove_all("\\)") %>%
    str_remove_all(",") ) %>%
  separate(gps_point, into = c("latitude", "longitude", "altitude", "accuracy"),
           sep = " ",
           convert = TRUE) %>%
  mutate(has_gps = !is.na(latitude) & !is.na(longitude)) %>% summary()
```

This seems to work, however not many households have GPS and there doesn't seem to be a lot
of information in the latitude and longitude values. Oh well...




This should suffice for the household head preprocessing. The function can be seen in action below:
```{r development-preprocess_2018_household_heads}
# Prepare the code of your function here
```

```{r function-preprocess_2018_household_heads}
#' preprocess_2018_household_heads Preprocess 2018 Household Heads
#'
#' @return Tibble with cleaned household head data
#' @export
preprocess_2018_household_heads <- function(opensrp, 
                                            cohort_2018) {
  
  hhreg <- opensrp$Hhregistration
  hhreg %>%
    janitor::clean_names() %>%
    mutate(hh_head_name = str_to_lower(name_household_head) %>% 
             str_squish() %>%
             str_remove_all(",")) %>%
    mutate(submission_date = ymd(submission_date)) %>%
    select(original_id = id, hh_head_name, submission_date, village, province, district) %>%
    mutate(across(where(is.character), as_factor)) -> hhreg_

  hh_geography <- opensrp$GPS
  hh_geography %>%
    janitor::clean_names() %>%
    mutate(hh_head_name = str_to_lower(head_of_household) %>% 
             str_squish() %>%
             str_remove_all("\\(") %>%
             str_remove_all("\\)") %>%
             str_remove_all(",")) %>%
    separate(gps_point, into = c("gps_latitude", "gps_longitude", "gps_altitude", "gps_accuracy"),
             sep = " ",
             convert = TRUE) %>%
    mutate(has_gps = !is.na(gps_latitude) & !is.na(gps_longitude)) %>%
    filter(has_gps) %>%
    select(hh_head_name, gps_latitude:gps_accuracy) %>%
    mutate(across(where(is.character), as_factor)) -> hh_geography_

  # Join
  hhreg_ %>%
    left_join(hh_geography_, by = "hh_head_name") %>%
    mutate(hh_head_name = case_when(
      hh_head_name == "kalo jsannette" ~ "kalo jeannette",
      hh_head_name == "emilien" ~ "emillien",
      hh_head_name == "randrianasoavana" ~ "randrianasoaviana",
      str_starts(original_id, "A3dd") ~ "eddy ii",
      TRUE ~ hh_head_name
    )) %>%
    left_join(
      cohort_2018 %>%
        select(uuid, name, hh_head_name, original_id),
    join_by(original_id == original_id, hh_head_name == name)) %>% select(-hh_head_name.y) %>%
    # there are two cases that come out with NA. In the first, it is `hervet`, which we can
    # assume was hervet testing the system. In the second, it is tsidinesana, which is an
    # actual duplicate in the data.
    filter(!is.na(uuid)) -> hhreg_final

  hhreg_final
}
```

```{r examples-preprocess_2018_household_heads}
preprocess_2018_household_heads(opensrp, cohort_2018) %>%
  summary()
```

```{r tests-preprocess_2018_household_heads}
test_that("preprocess_2018_household_heads works", {
  expect_true(inherits(preprocess_2018_household_heads, "function"))
})
```

```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_cohort_household_heads.Rmd", vignette_name = "Preprocessing 2018 Household Heads", check = FALSE)
```

