---
title: "Tracking Input Data"
author: "Tinashe M. Tapera"
output: rmarkdown::html_vignette
editor_options: 
  chunk_output_type: console
# vignette: >
#   %\VignetteIndexEntry{Tracking Input Data}
#   %\VignetteEngine{knitr::rmarkdown}
#   %\VignetteEncoding{UTF-8}
---

```{r development, include=FALSE}
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

# Tracking Inputs

In this notebook, we're developing functions to track all of the input data. Importantly,
this data is stored on Google Drive and must not be shared publicly on the internet,
so we will need to use programmatic credentials to access it.

Let's first create a function that will log us in to google drive with the appropriate credentials.
For this, we'll use the `googledrive` package from `tidyverse`.
```{r development-tracking_inputs}
# Prepare the code of your function here

# necessary libraries
library(googledrive)
```

We can start the process of authorization using the `drive_auth()` function. This requires
access to the a special JSON file that contains the credentials for the google drive account.

To enable non-interactive access to Google Drive from this R package, we use a _Google Cloud 
service account_ for authentication. This allows scripts and functions to interact with 
Google Drive without requiring manual sign-in or browser-based authentication. It’s particularly
useful for automated workflows, scheduled tasks, or when deploying the package in cloud 
environments (e.g., on a server like `FASRC`). The service account acts like
a dummy user with its own credentials, and it can be granted access to specific files or 
folders in the CSPH drives. Access to the service account can be controlled via sharing
permissions just like any human Google Drive user.

To set this up, the service account was created via the Google Cloud Console under
a project with the Google Drive API enabled. A JSON key was generated and 
downloaded — this file contains the credentials used to authenticate in R. 
Any Google Drive resources the package needs to access must be explicitly shared with 
the service account’s email (in this case, `data-pipeline@harvard-csph-driveauth.iam.gserviceaccount.com`).
Once shared, the pipeline uses the `googledrive` package in R to authenticate via 
`drive_auth(path = "path/to/key.json")`, allowing Drive operations to run seamlessly.

This authentication method ensures that all users of this package (or future maintainers)
can use a consistent, secured credential file for Drive access. To maintain this setup,
keep the JSON key file in a protected location, and do not commit it to version control. 
If the key ever needs to be regenerated (e.g., due to compromise or rotation), the same
process above can be followed to create a new one.

## Authenticating with Google Drive

If you have the authentication file, you can simply point to it with your `drive_auth()` function in
an interactive session. If this is successful, you should be able to access the Google Drive files
from R with `drive_find()`.


```{r, eval=FALSE}
path_to_auth <- "path/to/your/auth.json"
drive_auth(path = path_to_auth)
drive_find()
```

In a non-interactive session, make sure to set the environment variable with `.Renviron`.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(googledrive)
path_to_auth <- Sys.getenv("GOOGLE_DRIVE_AUTH_JSON")
drive_auth(path = path_to_auth)
drive_find() # does not print anything in vignettes but will print interactively
```

```{r function-authenticate_google_drive}
#' authenticate_google_drive Title
#'
#' Authenticate with Google Drive using the appropriate credentials
#' The pipeline will not run if this step is not successful.
#' 
#' @importFrom googledrive drive_auth
#' @return TRUE if the authentication was successful
#' @export
#'
#' @examples
#' authenticate_google_drive()
authenticate_google_drive <- function() {
  
  # Load the path to the authentication file
  path_to_auth <- Sys.getenv("GOOGLE_DRIVE_AUTH_JSON")

  tryCatch({
    
    drive_auth(path = path_to_auth)
    return(TRUE)

  }, error = function(e) {
    message("Authentication failed. Please check the path to the authentication file.")
    message(e)
    return(FALSE)
  })

}
```

```{r examples-authenticate_google_drive}
authenticate_google_drive()
```

```{r tests-tracking_inputs}
test_that("Google authentication works", {
  expect_true(authenticate_google_drive())
})
```

We can add this to the beginning of our `targets` pipeline.

```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_tracking_inputs.Rmd", vignette_name = "Tracking Input Data")
```

